---
title: "GAE7"
author: James Simon
format: 
  html:
    theme: custom.scss
    embed-resources: true
    code-tools: true

csl: ecology.csl

editor: visual
bibliography: references.bib
---

```{r}
#| label: SetUp
#| message: FALSE
#| echo: FALSE

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
  ) 

library(emmeans)
library(lmerTest)
library(effectsize)
library(car)
library(ggfortify)
library(patchwork)
library(broom)
library(latex2exp)       
library(kableExtra)
library(tidyverse)
library(redres)


options(knitr.kable.NA = '')
```

## Introduction

@Pintar.etal2018 conducted a field mesocosm experiment to explore the relative effects of predation risk and nutrient availability (a pdf of their publication is included in the GitHub repository). The predation risk manipulation was accomplished by adding either 0, 1, or 2, Golden Topminnows (*Fundulus chrysotus*) to each mesocosm. These small fish are surface-feeding predators and are known to elicit colonization and oviposition avoidance among dytiscid beetles and *Culex* mosquitoes, but not in some hydrophilid beetles. Nutrient-level treatments were created by the addition of 0, 4, or 8 g of rabbit chow (Small World Rabbit Food, Mannapro, St. Louis, MO, 40% protein).

## Methods

Please see @Pintar.etal2018 for greater detail on experimental design and data collection. In brief, the experiment was set up as a 3 X 3 factorial treatment design (9 treatment combinations). The experiment was replicated in 3 separate locations at 3 different times for a total of 9 replicates. This should have resulted in 81 observations total, however some fish died during the experiment rendering those observations not being included in the data resulting in only 69 observations.

```{r }
#| label: dataImport
#| tbl-cap: "First six observations of the data provided by @Pintar.etal2018"


df1 <- read_csv("Pintar_etal_2018.csv",
                show_col_types = FALSE)

# Block = 1 to 9 for the 9 spatiotemporal blocks (3 locations X 3 times)
# Pool = 1 to 9 for the plastic wading pools used for the experiment
# Location = 1 to 3 for the 3 fields where the pools were deployed
# Time = a to c for the 3 time periods used for each 2-week run of the exp.
# Nutrient = 1 to 3 for the 3 nutrient levels
# Fish = 1 to 3 for the 3 fish densities
# Culex_Egg_Rafts = number of egg rafts collected over the 2 weeks
# Hydrophilidae = number of hydrophilid beetles (Water Scavenger Beetles)
# Dytiscidae = number of dytiscid beetles (Predaceous Diving Beetles)

# Fish and Nutrient are provided as integers, but I wish to use them as ordered
# factors so that, by default, polynomial contrasts will be used to compute the
# ANOVAs.

df1$Nutrient <- factor(df1$Nutrient,
                       levels = c("0", "4", "8"),
                       ordered = TRUE)

df1$Fish <- factor(df1$Fish,
                   levels = c("0", "1", "2"),
                   ordered = TRUE)

df1$Block <- factor(df1$Block)

kbl(
  head(df1),
  col.names = c(
    "Block",
    "Pool",
    "Location",
    "Time",
    "Nutrient",
    "Fish",
    "Culex_Egg_Rafts",
    "Hydrophilidae",
    "Dytiscidae"
    )
  ) |>
  kable_classic(
    full = FALSE
    )


```

### Statistical Methods

One of the first steps we will take in order to analysize the data that is provided within @Pintar.etal2018 is to run an ANOVA summary on each of the variables we want to study. These are the number of Cullex Egg Rafts, Hydrophilidae, and Dytisciade. We want to see how related these are due to predation and the nutrient levels in the pool.

```{r}
#| label: tbl-fixedANOVA
#| tbl-cap: "ANOVA summary for the Hydrophilidae data provided by @Pintar.etal2018"
#| fig-show: hide

modFixed <- aov(log(0.5 + Hydrophilidae*10) ~ Pool + Block, data = df1)

plot(modFixed)

#| label: tbl-fixedANOVAFIN
#| tbl-cap: "ANOVA summary for the Hydrophilidae data provided by @Pintar.etal2018"
#| fig-show: hide

kbl(
  Anova(modFixed),
  digits = c(3,0,3,3)
  ) |>
  kable_classic()
```

```{r}
#| label: tbl-fixedANOVA2
#| tbl-cap: "ANOVA summary for the Cullex Egg Raft data provided by @Pintar.etal2018"
#| fig-show: hide

modFixed <- aov(log(0.5 + Culex_Egg_Rafts*10) ~ Pool + Block, data = df1)

plot(modFixed)

#| label: tbl-fixedANOVAFIN
#| tbl-cap: "ANOVA summary for the Cullex Egg Rafts data provided by @Pintar.etal2018"
#| fig-show: hide

kbl(
  Anova(modFixed),
  digits = c(3,0,3,3)
  ) |>
  kable_classic()

```

```{r}
#| label: tbl-fixedANOVA3
#| tbl-cap: "ANOVA summary for the Dytiscidae data provided by @Pintar.etal2018"
#| fig-show: hide

modFixed <- aov(log(0.5 + Dytiscidae*10) ~ Pool + Block, data = df1)

plot(modFixed)

#| label: tbl-fixedANOVAFIN
#| tbl-cap: "ANOVA summary for the Dytiscidae data provided by @Pintar.etal2018"
#| fig-show: hide

kbl(
  Anova(modFixed),
  digits = c(3,0,3,3)
  ) |>
  kable_classic()
```

After running a Fixed ANOVA, we can then make a linear model for each variable in order to see their correlation as well.

```{r}
#| label: tbl-HFactorial
#| tbl-cap: "Linear model summary for the Hydrophilidae  data provided by @Pintar.etal2018"
HFactorial <- lmer(Hydrophilidae ~ Fish*Nutrient + (1 | Block), data = df1)

HAnova <- (anova(HFactorial))

kbl(
  HAnova, col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "p"
    
  ),
  digits = c(3,3,0,0,3,4)
) |>
  kable_classic()

```

```{r}
#| label: tbl-CERFactorial
#| tbl-cap: "Linear model summary for the Cullex Egg Raft  data provided by @Pintar.etal2018"
CERFactorial <- lmer(Culex_Egg_Rafts ~ Fish*Nutrient + (1 | Block), data = df1)

CERAnova <- (anova(CERFactorial))

kbl(
  CERAnova, col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "p"
    
  ),
  digits = c(3,3,0,0,3,4)
) |>
  kable_classic()
```

```{r}
#| label: tbl-DyFactorial
#| tbl-cap: "Linear model summary for the Dytiscidae  data provided by @Pintar.etal2018"
DyFactorial <- lmer(Dytiscidae ~ Fish*Nutrient + (1 | Block), data = df1)

DyAnova <- (anova(DyFactorial))

kbl(
  DyAnova, col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "p"
    
  ),
  digits = c(3,3,0,0,3,4)
) |>
  kable_classic()
```

We can also conduct Mixed ANOVA Tests for the variables.

```{r}
#| label: tbl-HMixedAnova
#| tbl-cap: "Mixed-effect model results for Hydrophilidae and predation and nutrient levels"
#| tbl-subcap:
#|     - "Fixed effect"
#|     - "Random effect"
#| layout-nrow: 2

HMixed <- anova(HFactorial)

kbl(
  HMixed,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "p"
  ),
  digits = c(3,3,0,0,3,4)
) |>
  kable_classic()

kbl(
  rand(HFactorial),
  col.names = c(
    "Parameters",
    "log-liklehood",
    "AIC",
    "LRT",
    "df",
    "p"
  ),
  digits = c(0,3,3,3,0,4)
)|>
  kable_classic()
```

```{r}
#| label: tbl-CERMixedAnova
#| tbl-cap: "Mixed-effect model results for Cullex Egg Raft and predation and nutrient levels"
#| tbl-subcap:
#|     - "Fixed effect"
#|     - "Random effect"
#| layout-nrow: 2

CERMixed <- anova(CERFactorial)

kbl(
  CERMixed,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "p"
  ),
  digits = c(3,3,0,0,3,4)
) |>
  kable_classic()

kbl(
  rand(CERFactorial),
  col.names = c(
    "Parameters",
    "log-liklehood",
    "AIC",
    "LRT",
    "df",
    "p"
  ),
  digits = c(0,3,3,3,0,4)
)|>
  kable_classic()
```

```{r}
#| label: tbl-DyMixedAnova
#| tbl-cap: "Mixed-effect model results for Dytiscidae and predation and nutrient levels"
#| tbl-subcap:
#|     - "Fixed effect"
#|     - "Random effect"
#| layout-nrow: 2

DyMixed <- anova(DyFactorial)

kbl(
  DyMixed,
  col.names = c(
    "SS",
    "MS",
    "df1",
    "df2",
    "F",
    "p"
  ),
  digits = c(3,3,0,0,3,4)
) |>
  kable_classic()

kbl(
  rand(DyFactorial),
  col.names = c(
    "Parameters",
    "log-liklehood",
    "AIC",
    "LRT",
    "df",
    "p"
  ),
  digits = c(0,3,3,3,0,4)
)|>
  kable_classic()
```

## Results

Looking at the multiple tests, there are interesting trends that need to be discussed. Examining the Fixed ANOVA results, we can see that predation played an important role within controlling the rates of *Dytiscidae* as well as the rates of Cullex Egg Rafts.

## References
